State of the Practice for Medical Imaging Software

### 2022-04-01 StateOfPractice/Papers/MedImageSoft_SOP-arXiv/MedImageSoft_SOP_AcademicRadiology.tex
Updates to SOP for MI for the artifacts section

As part of filling in the measurement template the MI packages were examined for
the presence of artifacts, which were then categorized by frequency. We have
grouped them into common, less common, and rare artifacts in
Table~\ref{artifactspresent}. Common artifacts were found in 20 to 29 ($>$67\%)
of the software packages. Uncommon and rare artifacts were found in
10 to 19 (33-67\%) and 1 to 9 ($<$33\%) of the software packages, respectively.
The full measurements are summarized in \citet{Dong2021-Data}.  The details on
which projects use which types of artifacts is summarized in
Tables~\ref{tab_maintainability_docs} and~\ref{tab_Visibility/Transparency_docs}
for documents related to maintainability and visibility, respectively.

\begin{table}[ht!]
    \begin{center}
    \begin{tabular}{ p{4.6 cm} p{5.6 cm} p{5 cm}}
    \toprule
    \textbf{Common} & \textbf{Uncommon} & \textbf{Rare} \\
    \midrule
    README (29) & Build scripts (18) & Developer's manual (8)\\
    Version control (29) & Tutorials (18) & API documentation (7)\\
    License (28) & Installation guide (16) & Troubleshooting guide (6)\\
    Issue tracker (28) & Test cases (15) & Product roadmap (5)\\
    Change request (28) & Authors (14) & \\
    User manual (22) & Frequently Asked Questions (FAQ) (14) & \\
    Release info. (22) & Acknowledgements (12) & \\
    \bottomrule
    \end{tabular}
    \caption{Artifacts Present in MI Packages, Classified by Frequency (The number in brackets is the number of occurrences)}
    \label{artifactspresent}
    \end{center}
\end{table}

@@

Table~\ref{artifactspresent} shows that some types of software engineering
recommended artifacts are missing in MI projects.  
  A gap exists in scientific
computing development practices and software engineering recommendations
\citep{OwojaiyeEtAl2021_CSE}.

\begin{table}[!ht]
\begin{center}
\begin{tabular}{ p{2.5cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1.2cm}p{1cm}p{0.8cm} }
\toprule
~ \ & \cite{USGS2019} & \cite{TobiasEtAl2018} & \cite{BrettEtAl2021} & \cite{WilsonEtAl2016} & \cite{SmithAndRoscoe2018} & \cite{HerouxEtAl2008} & \cite{ThielEtAl2020} & \cite{vanGompelEtAl2016} & \cite{OrvizEtAl2017} & MI\\
\midrule
LICENSE & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & C\\
README &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & C\\
CONTRIBUTING &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark \\
CITATION &  &  &  & \checkmark & & & & \checkmark & \checkmark &  \\
CHANGELOG &  & \checkmark &  & \checkmark & \checkmark & & \checkmark &  &  &  \\
INSTALL &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Uninstall &  &  &  &  & & & & \checkmark & &  \\
Dependency List &  &  & \checkmark & & \checkmark & & & \checkmark &  &  \\
Authors &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code of Conduct &  &  &  &  & & & \checkmark & & &  \\
Acknowledgements &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code Style Guide &  & \checkmark &  &  & & & \checkmark & \checkmark & \checkmark &\\
Release Info. &  & \checkmark &  &  & & \checkmark & \checkmark & & & C\\
Prod.\ Roadmap &  &  &  &  & & \checkmark & \checkmark & \checkmark & & R\\
\midrule
Getting started &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark &  \\
User manual &  &  & \checkmark &  & & & \checkmark & & & C\\
Tutorials &  &  &  &  & & & \checkmark & & & U\\
FAQ &  &  &  &  & & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Issue Track &  & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & & \checkmark & C\\
Version Control &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & C\\ 
Build Scripts &  & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & U\\
\midrule
Requirements &  & \checkmark &  &  & & \checkmark &  &  & \checkmark &  \\
Design Doc.\ &  & \checkmark  & \checkmark &  & \checkmark & & \checkmark & \checkmark& \checkmark & \\
API Doc. &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
Test Plan &  & \checkmark &  &  & & \checkmark & & & &  \\
Test Cases & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & U\\
\bottomrule
\end{tabular}
\caption{Comparison of Recommended Artifacts in Software Development Guidelines
to Artifacts in MI Projects (C for Common, U for Uncommon and R for Rare)}
\label{Tbl_Guidelines}
    \end{center}
\end{table}

The majority of MI generated artifacts correspond to general recommendations
from research software developers.  A union of the three columns in
Table~\ref{artifactspresent} mostly corresponds to recommendations made by the
research software community.  Most of the items in Table~\ref{artifactspresent}
are included in guidelines for writing research software, such as the DLR
(German Aerospace Centre) Software Engineering Guidelines
\citep{TobiasEtAl2018}, xSDK (Extreme-scale Scientific Software Development Kit)
Community Package Policies \citep{SmithAndRoscoe2018}, the EURISE (European
Research Infrastructure Software Engineers') Network Technical Reference
\cite{ThielEtAl2020} and CLARIAH (Common Lab Research Infrastructure for the
Arts and Humanities) Guidelines for Software Quality \citep{vanGompelEtAl2016}.
As an example, MI developers commonly list the external libraries on which the
software depends, as recommended by \citet{BrettEtAl2021},
\citet{SmithAndRoscoe2018} and \citet{vanGompelEtAl2016}. Although in the rare
column, some MI developers include an FAQ (Frequently Asked Questions) document
as recommended by \citep{OrvizEtAl2017}, \citep{ThielEtAl2020} and
\citep{vanGompelEtAl2016}. MI developers also follow the ubiquitous
recommendation to have a README file, although, as for other software, the
content of README files is not standardized \citep{PranaEtAl2018}.  Compared to
ocean modelling software \citep{JungEtAl2022}, MI software pays more
attention to artifacts related to testing.

>>>>

As part of filling in the measurement template the MI packages were examined for
the presence of artifacts, which were then categorized by frequency.
Table~\ref{artifactspresent} groups them into categories of common (20 to 29
($>$67\%) packages), uncommon (10 to 19 (33-67\%) packages), and rare (1 to 9
($<$33\%) packages). The full measurements are summarized in
\citet{Dong2021-Data}.  The details on which projects use which types of
artifacts are summarized in Tables~\ref{tab_maintainability_docs}
and~\ref{tab_Visibility/Transparency_docs} for documents related to
maintainability and visibility, respectively.

\begin{table}[ht!]
    \begin{center}
    \begin{tabular}{ p{4.6 cm} p{5.6 cm} p{5 cm}}
    \toprule
    \textbf{Common} & \textbf{Uncommon} & \textbf{Rare} \\
    \midrule
    README (29) & Build scripts (18) & Getting Started (9)\\
    Version control (29) & Tutorials (18) & Developer's manual (8)\\
    License (28) & Installation guide (16) & Contributing (8)\\
    Issue tracker (28) & Test cases (15) & API documentation (7)\\
    Change request (28) & Authors (14) & Dependency List (7)\\
    User manual (22) & Frequently Asked Questions (FAQ) (14) & Troubleshooting guide (6)\\
    Release info. (22) & Acknowledgements (12) & Product roadmap (5)\\
     & Changelog (12) & Design Documentation (5)\\
     & Citation (11) & Code Style Guide (3)\\
     & & Code of Conduct (1)\\
    \bottomrule
    \end{tabular}
    \caption{Artifacts Present in MI Packages, Classified by Frequency (The number in brackets is the number of occurrences)}
    \label{artifactspresent}
    \end{center}
\end{table}

@@

Table~\ref{Tbl_Guidelines} shows that the MI artifacts generally match the recommendations found in nine current research software development guidelines:
\begin{inparaenum}[i)]
\item United States Geological Survey Software Planning Checklist
\citep{USGS2019},
\item DLR (German Aerospace Centre) Software Engineering Guidelines
\citep{TobiasEtAl2018}, 
\item Scottish Covid-19 Response Consortium Software Checklist
\citep{BrettEtAl2021},
\item Good Enough Practices in Scientific Computing \citep{WilsonEtAl2016},
\item xSDK (Extreme-scale Scientific Software Development Kit) Community Package
Policies \citep{SmithAndRoscoe2018},
\item Trilinos Developers Guide \citep{HerouxEtAl2008},
\item EURISE (European Research Infrastructure Software Engineers') Network
Technical Reference \citep{ThielEtAl2020},
\item CLARIAH (Common Lab Research Infrastructure for the Arts and Humanities)
Guidelines for Software Quality \citep{vanGompelEtAl2016}, and
\item A Set of Common Software Quality Assurance Baseline Criteria for Research
Projects \citep{OrvizEtAl2017}.
\end{inparaenum}
As examples of compliance with most guidelines, MI developers commonly include
the almost ubiquitous advice to have LICENSE files and README files. , although,
as for other software, the content of README files is not standardized
\citep{PranaEtAl2018}. 15 of 26 artifacts (58\%).  Although in the rare column, some MI developers include
an FAQ (Frequently Asked Questions) document as recommended by
\citep{OrvizEtAl2017}, \citep{ThielEtAl2020} and \citep{vanGompelEtAl2016}. MI
developers also follow the ubiquitous recommendation to have a README file
Compared to ocean modelling software \citep{JungEtAl2022}, MI software pays more
attention to artifacts related to testing.

\begin{table}[!ht]
\begin{center}
\begin{tabular}{ p{2.5cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1.2cm}p{1cm}p{0.8cm} }
\toprule
~ \ & \cite{USGS2019} & \cite{TobiasEtAl2018} & \cite{BrettEtAl2021} & \cite{WilsonEtAl2016} & \cite{SmithAndRoscoe2018} & \cite{HerouxEtAl2008} & \cite{ThielEtAl2020} & \cite{vanGompelEtAl2016} & \cite{OrvizEtAl2017} & MI\\
\midrule
LICENSE & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & C\\
README &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & C\\
CONTRIBUTING &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
CITATION &  &  &  & \checkmark & & & & \checkmark & \checkmark & U\\
CHANGELOG &  & \checkmark &  & \checkmark & \checkmark & & \checkmark &  &  & U\\
INSTALL &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Uninstall &  &  &  &  & & & & \checkmark & &  \\
Dependency List &  &  & \checkmark & & \checkmark & & & \checkmark &  & R\\
Authors &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code of Conduct &  &  &  &  & & & \checkmark & & & R\\
Acknowledgements &  &  &  &  &  &  & \checkmark & \checkmark & \checkmark & U\\
Code Style Guide &  & \checkmark &  &  & & & \checkmark & \checkmark & \checkmark & R\\
Release Info. &  & \checkmark &  &  & & \checkmark & \checkmark & & & C\\
Prod.\ Roadmap &  &  &  &  & & \checkmark & \checkmark & \checkmark & & R\\
\midrule
Getting started &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
User manual &  &  & \checkmark &  & & & \checkmark & & & C\\
Tutorials &  &  &  &  & & & \checkmark & & & U\\
FAQ &  &  &  &  & & & \checkmark & \checkmark & \checkmark & U\\
\midrule
Issue Track &  & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark & & \checkmark & C\\
Version Control &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & C\\ 
Build Scripts &  & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark & & \checkmark & U\\
\midrule
Requirements &  & \checkmark &  &  & & \checkmark &  &  & \checkmark &  \\
Design Doc.\ &  & \checkmark  & \checkmark &  & \checkmark & & \checkmark & \checkmark& \checkmark & R\\
API Doc. &  &  &  &  & \checkmark & & \checkmark & \checkmark & \checkmark & R\\
Test Plan &  & \checkmark &  &  & & \checkmark & & & &  \\
Test Cases & \checkmark & \checkmark & \checkmark &  & \checkmark & \checkmark & \checkmark & \checkmark & \checkmark & U\\
\bottomrule
\end{tabular}
\caption{Comparison of Recommended Artifacts in Software Development Guidelines
to Artifacts in MI Projects (C for Common, U for Uncommon and R for Rare)}
\label{Tbl_Guidelines}
\end{center}
\end{table}

A gap exists in scientific
computing development practices and software engineering recommendations
\citep{OwojaiyeEtAl2021_CSE}.
